---
title: "Assignment 4"
author: "Filippo Ravalli"
date: "2023-02-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Read data in & libraries
```{r}
library(tidyverse)
library(labelled)
library(caret)
library(cluster)
library(ggbiplot)
library(stats)
library(factoextra)
library(cluster)
```

## Part 1

## Data Cleaning

```{r}
data<-read.csv("class4_p1.csv")
glimpse(data)
data <- data %>%
  rename(hypten = chronic1, 
         diabetes = chronic3, 
         asthma=chronic4,
         cig_use = tobacco1,  
         alc_drink = alcohol1, 
         pa_min = gpaq8totmin, 
         walk_7days = gpaq11days, 
         physical_act_q = habits5,
         diet = habits7,
         gender=dem3, 
         hispanic=dem4,
         us_cit=dem8)

data <- data %>%
  set_value_labels(
    hypten = c("Yes" = 1, "No" = 2), 
    diabetes = c("Yes" = 1, "No" = 2), 
    asthma = c("Yes" = 1, "No" = 2), 
    pa_min = c("most days" = 1, "some days" = 2, "never" = 3), 
    physical_act_q = c("very active" = 1, "somewhat active" = 2, "not very active" = 3, "not active at all" = 4), 
    diet = c("excellent" = 1, "very good" = 2, "good"=3, "fair"=4, "poor"=5), 
    gender = c("male" = 1, "female" = 2), 
    hispanic = c("Yes" = 1, "No" = 2), 
    us_cit = c("USA" = 1, "outside USA"=2),
    cig_use = c("most days" = 1, "some days" = 2, "never" = 3), 
    alc_drink = c("most days" = 1, "some days" = 2, "never" = 3), 
    agegroup=c("18-24"=1, "25-44"=2, "45-64"=3, "65+"=4),
    povertygroup=c("<100%"=1,"100-99%"=2, "200-399%"=3, "400-599%"=4, "600%+"=5, "dont know"=6))
    
      

data <- data %>%
  mutate_if(is.labelled, to_factor)
glimpse(data)
```
**Data that are continuous are bmi, walk7days (gpaq11days), & healthydays. The other variables are categorical.**

## Data Partition
```{r}
set.seed(123)
data_na_omit<-na.omit(data)
train_index<-createDataPartition(data_na_omit$healthydays, p=0.7, list=FALSE)

data_train<-data_na_omit[train_index,]
data_test<-data_na_omit[-train_index,]
```

## Fit prediction models
```{r}
model1<-lm(healthydays~bmi+diabetes+agegroup, data=data_train)
summary(model1)
model2<-lm(healthydays~gender+povertygroup+us_cit+hispanic, data=data_train)
summary(model2)
```

## Test models
```{r}
set.seed(123)
predict_model1<-model1 %>% predict(data_test)
RMSE(predict_model1, data_test$healthydays) #7.463758

predict_model2<-model2 %>% predict(data_test)
RMSE(predict_model2, data_test$healthydays) #7.540379
```

**Question 2**
**Based on the RMSE values, model 1 is the more accurate as it has a lower RMSE comapred to model 2.**

**Question 3**
**The implementation of the final model would be helpful in a research setting where the question is trying to understand the effects of social factors, including Hispanic status, place of birth, poverty group, and gender.**


## Part 2
## Hierarchical analysis

```{r}
arrests<-USArrests

arrests_nomiss<-na.omit(arrests)

colMeans(arrests_nomiss, na.rm=TRUE)
apply(arrests_nomiss, 2, sd, na.rm=TRUE)


# Create Dissimilarity matrix
diss.matrix <- dist(arrests_nomiss, method = "euclidean")

# Hierarchical clustering using Complete Linkage
clusters.h<- hclust(diss.matrix, method = "complete" )

# Plot the obtained dendrogram
plot(clusters.h, cex = 0.6, hang = -1)

#create function to use within clusGap
hclusCut <- function(x, k) list(cluster = cutree(hclust(dist(x, method="euclidian"), method="average"), k=k))

gap_stat <- clusGap(arrests_nomiss, FUN = hclusCut, K.max = 10, B = 50)
fviz_gap_stat(gap_stat)

clusters.h.3<-cutree(clusters.h, k=3)
table(clusters.h.3)

#Alternatives for hierarchical clustering

clusters.hcut<-hcut(arrests_nomiss, k=5, hc_func="hclust", hc_method="single", hc_metric="euclidian")

clusters.hcut$size
fviz_dend(clusters.hcut, rect=TRUE)
fviz_cluster(clusters.hcut)

gap_stat <- clusGap(arrests_nomiss, FUN = hcut, hc_method="single", K.max = 10, B = 5)
fviz_gap_stat(gap_stat)

input.feature.vals<-cbind(arrests_nomiss,cluster=clusters.hcut$cluster)

input.feature.vals %>%
  group_by(cluster) %>%
  summarise_all(mean)
```

##

##

